Wikipedia WALS Feature Analysis

### 2019-07-17 surprisal/WALS_features.Rmd
plotting feature space vs cosine space distance

>>>>

```{r feature space}
imputed_df <- read_csv(here("Data/imputed_wals.csv"))

feature_space <- imputed_df %>% 
  gather(feature, value, -language) %>% 
  group_by(feature) %>% 
  mutate(value = as.numeric(as.factor(value))) %>% 
  spread(feature, value) %>%
  column_to_rownames("language") %>%
  dist(method = "manhattan") %>%
  as.matrix() %>%
  melt(value.name = "feature") %>%
  as_tibble() %>%
  rename(language1 = Var1, language2 = Var2) 
```

```{r cosines}
unigram_cosines <- wiki %>%
  filter(language %in% common_languages, gram == "Unigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, unigram_cosine, -language1) %>%
  mutate(unigram_cosine = 1 - abs(unigram_cosine))

bigram_cosines <- wiki %>%
  filter(language %in% common_languages, gram == "Bigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, bigram_cosine, -language1) %>%
  mutate(bigram_cosine = 1 - abs(bigram_cosine))

trigram_cosines <- wiki %>%
  filter(language %in% common_languages, gram == "Trigram") %>%
  select(-gram) %>% 
  column_to_rownames("language") %>%
  t() %>%
  cosine() %>%
  as_data_frame(rownames = "language1") %>%
  gather(language2, trigram_cosine, -language1) %>%
  mutate(trigram_cosine = 1 - abs(trigram_cosine))

all_distances <- feature_space %>%
  left_join(unigram_cosines) %>%
  left_join(bigram_cosines) %>%
  left_join(trigram_cosines)

all_distances %>% 
  ggplot(aes(x = feature)) +
    geom_smooth(aes(y = unigram_cosine, color = "#0033CC"), 
                method = 'lm', se = F) + 
    geom_smooth(aes(y = bigram_cosine, color = "#6699FF"), 
                method = 'lm', se = F) +
    geom_smooth(aes(y = trigram_cosine, color = "#33CCFF"), 
                method = 'lm', se = F) +
    ylab("Cosine Space Distance") + 
    xlab("Feature Space Distance") + 
    scale_color_manual(scale_name = "gram", 
                       values = c("#0033CC", "#6699FF", "#33CCFF"), 
                       labels = c("Unigrams", "Bigrams", "Trigrams")) + 
    ggtitle("Cosine space distance regressed on feature spaced distance")
```
